---
title: "Case study: FMEA risk on a pilot submission"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case study: FMEA risk on a pilot submission}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A risk log that only lists risks is a to-do list. What a reviewer wants to see
is that the risks were ranked, worked, and reduced. `r4subrisk` puts numbers on
that story using a Failure Modes and Effects Analysis (FMEA) model: every risk
gets a probability, an impact, and a detectability score, and their product is
the Risk Priority Number (RPN).

This case study runs the example submission risk register from `r4subdata`
through that model, then shows what happens when the team starts mitigating.

```{r libs, message = FALSE}
library(r4subrisk)
library(r4subdata)
```

## The register

`risk_register_pharma` is a worked submission risk log: 18 risks across data
quality, traceability, and documentation, each already scored on the three FMEA
dimensions.

```{r register}
risk_register_pharma[, c("risk_id", "category", "probability", "impact",
                         "detectability")]
```

`create_risk_register()` validates the scales and computes the RPN and risk
level for each row.

```{r build}
rr <- create_risk_register(risk_register_pharma)
rr
```

## The risk picture

`compute_risk_scores()` rolls the register up: the mean and maximum RPN, an
overall 0 to 1 risk score, and the spread of risk levels.

```{r scores}
scores <- compute_risk_scores(rr)
scores
scores$category_summary
```

The category summary is where prioritisation starts. The categories carrying the
highest mean RPN are the ones to work first.

## Working the highest risks

Mitigation is the point of the exercise. Suppose the team addresses the two
highest-RPN risks: they add the missing SDTM variables and document the ADaM
derivation, which lowers both the probability and the detectability difficulty.

```{r ranked}
ranked <- rr[order(-rr$rpn), c("risk_id", "description", "rpn", "risk_level")]
head(ranked, 3)
```

`apply_mitigations()` takes the updates and recomputes the RPNs and levels.

```{r mitigate}
top_two <- head(ranked$risk_id, 2)

updates <- data.frame(
  risk_id       = top_two,
  probability   = c(1, 2),
  detectability = c(1, 1),
  status        = "mitigated",
  stringsAsFactors = FALSE
)

rr_after <- apply_mitigations(rr, updates)
```

## Before and after

`compare_risk_registers()` reports what moved: which RPNs dropped, which risk
levels changed, and the shift in the mean.

```{r compare}
comparison <- compare_risk_registers(rr, rr_after)
comparison$rpn_changes[comparison$rpn_changes$rpn_delta != 0, ]
comparison$delta_mean_rpn
```

The mean RPN falls and the two worked risks drop to lower levels. That is the
before-and-after a reviewer is looking for: not that the study had no risks, but
that the risks were found, ranked, and brought down.

## Feeding the readiness score

The risk register does not live in isolation. `risk_register_to_evidence()`
turns it into R4SUB evidence rows, which flow into the Risk pillar of the
Submission Confidence Index alongside quality, trace, and usability evidence.
The FMEA work you do here shows up directly in the submission's readiness score.
