---
title: "Risk Management with r4subrisk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Risk Management with r4subrisk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Clinical submissions require a systematic approach to identifying, scoring,
and mitigating risks. The `r4subrisk` package implements an FMEA (Failure
Mode and Effects Analysis) based risk framework aligned with ICH Q9 quality
risk management principles.

```{r load}
library(r4subrisk)
```

## Default configuration

`risk_config_default()` defines the RPN (Risk Priority Number) band
thresholds and scoring scales:

```{r config}
cfg <- risk_config_default()
cfg$rpn_bands
cfg$required_origins
```

## Creating a risk register

A risk register is built from a data frame with at minimum `risk_id`,
`description`, `probability`, `impact`, and `detectability` — each scored
1–5. The RPN is computed as `probability * impact * detectability`.

```{r risk-register}
risks <- data.frame(
  risk_id       = paste0("RISK-00", 1:5),
  description   = c(
    "Missing required variables in ADSL",
    "Incomplete derivation documentation for ADTTE",
    "Independent QC not completed for ADRS",
    "Define-XML missing codelist references",
    "ADRG not included in submission package"
  ),
  category      = c("data_quality", "traceability", "process",
                    "documentation", "regulatory"),
  probability   = c(3L, 4L, 2L, 2L, 1L),
  impact        = c(4L, 5L, 5L, 3L, 5L),
  detectability = c(2L, 3L, 4L, 2L, 1L),
  owner         = c("Data Mgmt", "Stats", "Stats", "Programming", "Regulatory"),
  stringsAsFactors = FALSE
)

rr <- create_risk_register(risks)
print(rr)
```

## Computing risk scores

`compute_risk_scores()` returns aggregate metrics including overall risk
score (normalized 0-1), mean RPN, and per-category breakdown:

```{r risk-scores}
rs <- compute_risk_scores(rr)
rs$overall_risk_score
rs$mean_rpn
rs$risk_distribution
rs$category_summary
```

## Applying mitigations

`apply_mitigations()` updates the register with post-control scores to show
residual risk after mitigations are applied:

```{r mitigations}
controls <- data.frame(
  risk_id               = c("RISK-001", "RISK-002", "RISK-003"),
  mitigation            = c(
    "Automated variable check implemented in CI pipeline",
    "Derivation review added to QC checklist",
    "QC sign-off required before dataset lock"
  ),
  probability_residual  = c(1L, 2L, 1L),
  impact_residual       = c(4L, 5L, 5L),
  detectability_residual = c(1L, 1L, 1L),
  stringsAsFactors = FALSE
)

rr_mitigated <- apply_mitigations(rr, controls)
```

## Comparing registers

`compare_risk_registers()` shows the RPN change before and after
mitigations:

```{r compare}
comparison <- compare_risk_registers(rr, rr_mitigated)
comparison
```

## Integration with r4subcore evidence

`evidence_to_risks()` converts an evidence table with `indicator_domain == "risk"`
into a risk register format, allowing automation of risk identification
from evidence outputs:

```{r evidence-to-risk, eval=FALSE}
# Requires r4subcore
ctx <- r4subcore::r4sub_run_context(study_id = "STUDY01")
ev <- r4subcore::as_evidence(
  data.frame(
    asset_type = "program", asset_id = "prod_adrs.R",
    source_name = "r4subrisk", source_version = "0.1.0",
    indicator_id = "R-001", indicator_name = "Program Validation",
    indicator_domain = "risk", severity = "high",
    result = "fail", metric_value = 0, metric_unit = "score",
    message = "DVP not submitted", location = "prod_adrs.R",
    evidence_payload = "{}", stringsAsFactors = FALSE
  ), ctx = ctx
)
risk_df <- evidence_to_risks(ev)
rr_from_evidence <- create_risk_register(risk_df)
```
